Comb associative operads

### 2018-09-01 CAs_operads.tex
partie 3.3

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The \texorpdfstring{$3$}{3}-comb associative operad}
\label{subsec:CAs_3}
We now focus on the study of the operad $\CAs{3}$. By definition, this
operad is the quotient of $\Mag$ by the operad congruence spanned by the
relation
\begin{equation} \label{equ:rew_1}
    \begin{tikzpicture}[xscale=.22,yscale=.23,Centering]
        \node(0)at(0.00,-5.25){};
        \node(2)at(2.00,-5.25){};
        \node(4)at(4.00,-3.50){};
        \node(6)at(6.00,-1.75){};
        \node[NodeST](1)at(1.00,-3.50){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-1.75){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,0.00){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(3);
        \draw[Edge](2)--(1);
        \draw[Edge](3)--(5);
        \draw[Edge](4)--(3);
        \draw[Edge](6)--(5);
        \node(r)at(5.00,1.5){};
        \draw[Edge](r)--(5);
    \end{tikzpicture}
    \enspace \Rew \enspace
    \begin{tikzpicture}[xscale=.22,yscale=.23,Centering]
        \node(0)at(0.00,-1.75){};
        \node(2)at(2.00,-3.50){};
        \node(4)at(4.00,-5.25){};
        \node(6)at(6.00,-5.25){};
        \node[NodeST](1)at(1.00,0.00){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-1.75){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,-3.50){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](2)--(3);
        \draw[Edge](3)--(1);
        \draw[Edge](4)--(5);
        \draw[Edge](5)--(3);
        \draw[Edge](6)--(5);
        \node(r)at(1.00,1.5){};
        \draw[Edge](r)--(1);
    \end{tikzpicture}\,.
\end{equation}
This rewrite rule is compatible with the lexicographic order on prefix
words presented at the beginning of Section~\ref{sec:operad_Mag} in the
sense that the prefix word of the left member of~\eqref{equ:rew_1} is
lexicographically greater than the prefix word of the right one.
\medbreak

However, the rewrite relation $\RewContext$ induced by $\Rew$ is not
confluent. Indeed, we have
\begin{equation} \label{equ:branching_pair_CAs_3}
    \begin{tikzpicture}[xscale=.22,yscale=.22,Centering]
        \node(0)at(0.00,-7.20){};
        \node(2)at(2.00,-7.20){};
        \node(4)at(4.00,-5.40){};
        \node(6)at(6.00,-3.60){};
        \node(8)at(8.00,-1.80){};
        \node[NodeST](1)at(1.00,-5.40){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-3.60){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,-1.80){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,0.00){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(3);
        \draw[Edge](2)--(1);
        \draw[Edge](3)--(5);
        \draw[Edge](4)--(3);
        \draw[Edge](5)--(7);
        \draw[Edge](6)--(5);
        \draw[Edge](8)--(7);
        \node(r)at(7.00,1.35){};
        \draw[Edge](r)--(7);
    \end{tikzpicture}
    \enspace \RewContext \enspace
    \begin{tikzpicture}[xscale=.22,yscale=.20,Centering]
        \node(0)at(0.00,-4.50){};
        \node(2)at(2.00,-4.50){};
        \node(4)at(4.00,-4.50){};
        \node(6)at(6.00,-6.75){};
        \node(8)at(8.00,-6.75){};
        \node[NodeST](1)at(1.00,-2.25){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,0.00){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,-2.25){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,-4.50){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(3);
        \draw[Edge](2)--(1);
        \draw[Edge](4)--(5);
        \draw[Edge](5)--(3);
        \draw[Edge](6)--(7);
        \draw[Edge](7)--(5);
        \draw[Edge](8)--(7);
        \node(r)at(3.00,1.74){};
        \draw[Edge](r)--(3);
    \end{tikzpicture}
    \qquad \mbox{and} \qquad
    \begin{tikzpicture}[xscale=.22,yscale=.22,Centering]
        \node(0)at(0.00,-7.20){};
        \node(2)at(2.00,-7.20){};
        \node(4)at(4.00,-5.40){};
        \node(6)at(6.00,-3.60){};
        \node(8)at(8.00,-1.80){};
        \node[NodeST](1)at(1.00,-5.40){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-3.60){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,-1.80){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,0.00){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(3);
        \draw[Edge](2)--(1);
        \draw[Edge](3)--(5);
        \draw[Edge](4)--(3);
        \draw[Edge](5)--(7);
        \draw[Edge](6)--(5);
        \draw[Edge](8)--(7);
        \node(r)at(7.00,1.35){};
        \draw[Edge](r)--(7);
    \end{tikzpicture}
    \enspace \RewContext \enspace
    \begin{tikzpicture}[xscale=.22,yscale=.22,Centering]
        \node(0)at(0.00,-3.60){};
        \node(2)at(2.00,-5.40){};
        \node(4)at(4.00,-7.20){};
        \node(6)at(6.00,-7.20){};
        \node(8)at(8.00,-1.80){};
        \node[NodeST](1)at(1.00,-1.80){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-3.60){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,-5.40){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,0.00){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(7);
        \draw[Edge](2)--(3);
        \draw[Edge](3)--(1);
        \draw[Edge](4)--(5);
        \draw[Edge](5)--(3);
        \draw[Edge](6)--(5);
        \draw[Edge](8)--(7);
        \node(r)at(7.00,1.5){};
        \draw[Edge](r)--(7);
    \end{tikzpicture}\,,
\end{equation}
and the two right members of~\eqref{equ:branching_pair_CAs_3} form a
branching pair which is not joinable.
\medbreak

In order to transform the rewrite relation induced by~\eqref{equ:rew_1}
into a convergent one, we apply the Buchberger algorithm for
operads~\cite[Section 3.7]{DK10} with respect to the lexicographic order
on prefix words. Following this algorithm, we need to put the right
members of~\eqref{equ:branching_pair_CAs_3} in relation by $\Rew$. To
respect the lexicographic property of the prefix words, this leads to
the new relation
\begin{equation} \label{equ:rew_2}
    \begin{tikzpicture}[xscale=.22,yscale=.22,Centering]
        \node(0)at(0.00,-3.60){};
        \node(2)at(2.00,-5.40){};
        \node(4)at(4.00,-7.20){};
        \node(6)at(6.00,-7.20){};
        \node(8)at(8.00,-1.80){};
        \node[NodeST](1)at(1.00,-1.80){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-3.60){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,-5.40){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,0.00){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(7);
        \draw[Edge](2)--(3);
        \draw[Edge](3)--(1);
        \draw[Edge](4)--(5);
        \draw[Edge](5)--(3);
        \draw[Edge](6)--(5);
        \draw[Edge](8)--(7);
        \node(r)at(7.00,1.5){};
        \draw[Edge](r)--(7);
    \end{tikzpicture}
    \enspace \Rew \enspace
    \begin{tikzpicture}[xscale=.22,yscale=.20,Centering]
        \node(0)at(0.00,-4.50){};
        \node(2)at(2.00,-4.50){};
        \node(4)at(4.00,-4.50){};
        \node(6)at(6.00,-6.75){};
        \node(8)at(8.00,-6.75){};
        \node[NodeST](1)at(1.00,-2.25){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,0.00){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,-2.25){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,-4.50){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(3);
        \draw[Edge](2)--(1);
        \draw[Edge](4)--(5);
        \draw[Edge](5)--(3);
        \draw[Edge](6)--(7);
        \draw[Edge](7)--(5);
        \draw[Edge](8)--(7);
        \node(r)at(3.00,1.74){};
        \draw[Edge](r)--(3);
    \end{tikzpicture}\,.
\end{equation}
The Buchberger algorithm applied on binary trees of degrees $5$, $6$,
and $7$ provides the new relations \smallbreak
\input{Buchberger_relations_5-7}
\medbreak

@@

\Todo{S: résultats expérimentaux de quelques $\CAs{\gamma}$ à donner
dans cette partie (ou peut être dans une sous-partie précédente ; à
voir selon ce que ça donne).}

%By computer exploration, we have already some information in this
%direction: we have the sequence
%\begin{equation}
    %1, 1, 2, 5, 13, 35, 96, 264, 724, 1973, 5335, 14390, 38872, 105141,
    %284929, 774254
%\end{equation}
%for the first dimensions for $\CAs{4}$. By applying the Buchberger
%algorithm on trees of degrees until $10$, we obtain that a convergent
%orientation of $\CongrCAs{4}$ has, arity by arity, the sequence
%\begin{math}
    %0, 0, 0, 0, 1, 1, 0, 3, 4, 5, 18, 22
%\end{math}
%for its first cardinalities. Moreover, for $\CAs{5}$, we get the
%sequence
%\begin{equation}
%1, 1, 2, 5, 14, 41, 124, 384, 1210, 3861, 12440, 40392, 131997,
%433782, 1432696, 4752857
%\end{equation}
%of dimensions and the first cardinalities
%\begin{math}
    %0, 0, 0, 0, 0, 1, 1, 0, 0, 4, 5
%\end{math}
%for any convergent orientation of $\CongrCAs{5}$. Finally, for
%$\CAs{6}$, we get the sequence
%\begin{equation}
   %1, 1, 2, 5, 14, 42, 131, 420, 1375, 4576, 15431, 52598,
   %180895, 626862, 2186504, 7670138
%\end{equation}
%of dimensions and the first cardinalities
%\begin{math}
    %0, 0, 0, 0, 0, 0, 1, 1, 0, 0, 0
%\end{math}
%for any convergent orientation of $\CongrCAs{6}$. We can notice that
%only $\CAs{3}$ seems to have oscillating first dimensions.
%\medbreak

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Completion of comb associative operads}
We are now looking for finite convergent presentation of comb associative
 operads.
By definition, the operad $\CAs{\gamma}$ is the quotient of $\Mag$ by the 
operad congruence spanned by the relation
\begin{equation} \label{equ:rew_1}
    \LComb{\gamma}
    \enspace \Rew \enspace
    \RComb{\gamma}\,.
\end{equation}
This rewrite rule is compatible with the lexicographic order on prefix
words presented at the beginning of Section~\ref{sec:operad_Mag} in the
sense that the prefix word of the left member of~\eqref{equ:rew_1} is
lexicographically greater than the prefix word of the right one.
\medbreak

However, the rewrite relation $\RewContext$ induced by $\Rew$ is not
confluent. Indeed, we have
\begin{equation} \label{equ:branching_pair_CAs_3}
    \LComb{\gamma+1}
    \enspace \RewContext \enspace
    \begin{tikzpicture}[xscale=.22,yscale=.20,Centering]
        \node(0)at(0.00,-4.50){};
        \node(2)at(2.00,-4.50){};
        \node(6)at(6.00,-6.75){};
        \node(8)at(8.00,-6.75){};
        \node[NodeST](1)at(1.00,-2.25){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,0.00){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.30,-1.4){\begin{math}\gamma\end{math}};
        \node[NodeST](7)at(7.00,-4.50){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(3);
        \draw[Edge](2)--(1);
        \draw[Edge](6)--(7);
        \draw[Edge, dashed](7)--(3);
        \draw[Edge](8)--(7);
        \node(r)at(3.00,1.74){};
        \draw[Edge](r)--(3);
    \end{tikzpicture}
    \qquad \mbox{and} \qquad
    \LComb{\gamma+1}
    \enspace \RewContext \enspace
    \begin{tikzpicture}[xscale=.22,yscale=.22,Centering]
        \node(0)at(0.00,-3.60){};
        \node(4)at(4.00,-7.20){};
        \node(6)at(6.00,-7.20){};
        \node(8)at(8.00,-1.80){};
        \node[NodeST](1)at(1.00,-1.80){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.25,-2.90){\begin{math}\gamma\end{math}};
        \node[NodeST](5)at(5.00,-5.40){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,0.00){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(7);
        \draw[Edge,dashed](5)--(1);
        \draw[Edge](4)--(5);
        \draw[Edge](6)--(5);
        \draw[Edge](8)--(7);
        \node(r)at(7.00,1.5){};
        \draw[Edge](r)--(7);
    \end{tikzpicture}\,,
\end{equation}
and the two right members of~\eqref{equ:branching_pair_CAs_3} form a
branching pair which is not joinable.
\medbreak

In order to transform the rewrite relation induced by~\eqref{equ:rew_1}
into a convergent one, we apply the Buchberger algorithm for
operads~\cite[Section 3.7]{DK10} with respect to the lexicographic order
on prefix words. We first focus on the special case $\gamma = 3$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{The \texorpdfstring{$3$}{3}-comb associative operad}
\label{subsubsec:CAs_3}
Following the Buchberger algorithm, we need to put the right
members of~\eqref{equ:branching_pair_CAs_3} in relation by $\Rew$. To
respect the lexicographic property of the prefix words, this leads to
the new relation
\begin{equation} \label{equ:rew_2}
    \begin{tikzpicture}[xscale=.22,yscale=.22,Centering]
        \node(0)at(0.00,-3.60){};
        \node(2)at(2.00,-5.40){};
        \node(4)at(4.00,-7.20){};
        \node(6)at(6.00,-7.20){};
        \node(8)at(8.00,-1.80){};
        \node[NodeST](1)at(1.00,-1.80){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-3.60){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,-5.40){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,0.00){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(7);
        \draw[Edge](2)--(3);
        \draw[Edge](3)--(1);
        \draw[Edge](4)--(5);
        \draw[Edge](5)--(3);
        \draw[Edge](6)--(5);
        \draw[Edge](8)--(7);
        \node(r)at(7.00,1.5){};
        \draw[Edge](r)--(7);
    \end{tikzpicture}
    \enspace \Rew \enspace
    \begin{tikzpicture}[xscale=.22,yscale=.20,Centering]
        \node(0)at(0.00,-4.50){};
        \node(2)at(2.00,-4.50){};
        \node(4)at(4.00,-4.50){};
        \node(6)at(6.00,-6.75){};
        \node(8)at(8.00,-6.75){};
        \node[NodeST](1)at(1.00,-2.25){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,0.00){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,-2.25){\begin{math}\Product\end{math}};
        \node[NodeST](7)at(7.00,-4.50){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](1)--(3);
        \draw[Edge](2)--(1);
        \draw[Edge](4)--(5);
        \draw[Edge](5)--(3);
        \draw[Edge](6)--(7);
        \draw[Edge](7)--(5);
        \draw[Edge](8)--(7);
        \node(r)at(3.00,1.74){};
        \draw[Edge](r)--(3);
    \end{tikzpicture}\,.
\end{equation}
The Buchberger algorithm applied on binary trees of degrees $5$, $6$,
and $7$ provides the new relations \smallbreak
\input{Buchberger_relations_5-7}
\medbreak

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Higher comb asociative operads}

We run the same algorithm for $\CAs{\gamma}$ when $\gamma \geq 4$.
We obtain the following results.

\begin{center}
  \begin{math}
    \begin{array}{|c|l|}
      \hline
      \gamma & \mbox{The sequence of first cardinalities of } 
       \CAs{\gamma} \mbox{completion}\\ \hline
      3 & 0, 0, 0, 1, 1, 2, 3, 4, 0, 0, \dots \\ \hline
      4 & 0, 0, 0, 0, 1, 1, 0, 3, 4, 5, 18, 22, 11, 12, 15, 19, 25, 36,
       44, 52, 68, 79, 93, 105, 106, 109 \\ \hline
      5 & 0, 0, 0, 0, 0, 1, 1, 0, 0, 4, 5, 8, 18, 31, 36, 48, 73, 111,
       172, 272, 455, 783 \\ \hline
      6 & 0, 0, 0, 0, 0, 0, 1, 1, 0, 0, 0, 5, 6, 11, 23, 30, 48, 73, 
      117, 204, 348, 589, 1004 \\ \hline
      7 & 0, 0, 0, 0, 0, 0, 0, 1, 1, 0, 0, 0, 0, 6, 7, 16, 24, 32, 49, 
      88, 150, 261, 475, 854\\ \hline
      8 & 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 0, 0, 0, 0, 0, 7, 8, 21, 29, 
      34, 53, 93, 172, 311, 565\\ \hline
      9 & 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 0, 0, 0, 0, 0, 0, 8, 9, 28, 
      30, 36, 57, 101, 185, 348, 648\\ \hline
    \end{array}
  \end{math}
\end{center}

From these computer explorations, we conjecture that this algorithm
doesn't provide a finite convergent presentation of $\CAs{\gamma}$, 
when $\gamma \geq 4$.

However, this completion algorithm depends on the other we choose. 
We might found a finite presentation using a different order.

\Todo{Expliquer algo et résultats}

We conjecture that there is no finite and convergent presentation
of $\CAs{\gamma}$ with one generator when $\gamma \geq 4$.
We discuss the cases with more then one generator in the perspectives.

%By computer exploration, we have already some information in this
%direction: we have the sequence
%\begin{equation}
    %1, 1, 2, 5, 13, 35, 96, 264, 724, 1973, 5335, 14390, 38872, 105141,
    %284929, 774254
%\end{equation}
%for the first dimensions for $\CAs{4}$. Moreover, for $\CAs{5}$, 
%we get the
%sequence
%\begin{equation}
%1, 1, 2, 5, 14, 41, 124, 384, 1210, 3861, 12440, 40392, 131997,
%433782, 1432696, 4752857
%\end{equation}
%of dimensions and the first cardinalities
%\begin{math}
    %0, 0, 0, 0, 0, 1, 1, 0, 0, 4, 5
%\end{math}
%for any convergent orientation of $\CongrCAs{5}$. Finally, for
%$\CAs{6}$, we get the sequence
%\begin{equation}
   %1, 1, 2, 5, 14, 42, 131, 420, 1375, 4576, 15431, 52598,
   %180895, 626862, 2186504, 7670138
%\end{equation}
%of dimensions and the first cardinalities
%\begin{math}
    %0, 0, 0, 0, 0, 0, 1, 1, 0, 0, 0
%\end{math}
%for any convergent orientation of $\CongrCAs{6}$. We can notice that
%only $\CAs{3}$ seems to have oscillating first dimensions.
%\medbreak
